Efficient Thresholded Correlation using Truncated Singular Value Decomposition

### 2015-07-22 threshold/draft.Rnw
notes

Consider a real-valued data matrix $A$ consisting of $m$ observations of $n$
column vectors.  Denote the columns of $A$ as $a_j\in\mathcal{R}^m$ for
$j=1,2,\ldots,n$, $A=[a_1, a_2, \ldots, a_n]\in\mathcal{R}^{m\times n}$, and
let $k=\mathrm{rank}(A)$.  Assume that the mean of each column $a_j$ is zero
and that each column has unit norm $\|a_j\|^2 = 1$. Here and below, $\|\cdot\|$
denotes the Euclidean vector norm. Then the sample correlation matrix
$\mathrm{cor}(A)=A^TA$.  Note that under these assumptions the sample
correlation and covariance matrices are the same.  Section \ref{irlba} below
illustrates how to relax the unit norm and zero mean assumptions in practice.

@@

We have so far assumed that the mean of each column vector of $A$ is zero and
the columns are scaled to have norm 1, but we really want a method that works
with general matrices $A$. We've also implicitly assumed that we can
efficiently compute a relatively low-rank truncated SVD of potentially
large  matrices required by step 1 of Algorithm \ref{prune}.  Fortunately,
both needs are met by one algorithm.

>>>>

Consider a real-valued data matrix $A$ consisting of $m$ observations of $n$
column vectors.  Denote the columns of $A$ as $a_j\in\mathcal{R}^m$ for
$j=1,2,\ldots,n$, $A=[a_1, a_2, \ldots, a_n]\in\mathcal{R}^{m\times n}$, and
let $k=\mathrm{rank}(A)$.  Assume that the mean of each column $a_j$ is zero
and that each column has unit norm $\|a_j\| = 1$. Here and below, $\|\cdot\|$
denotes the Euclidean vector norm. Then the sample correlation matrix
$\mathrm{cor}(A)=A^TA$.  Note that under these assumptions the sample
correlation and covariance matrices are the same.  Section \ref{irlba} below
illustrates how to relax the unit norm and zero mean assumptions in practice.

@@

We have so far assumed that the mean of each column vector of $A$ is zero and
the columns are scaled to have unit norm, but we really want a method that works
with general matrices $A$. We've also implicitly assumed that we can
efficiently compute a relatively low-rank truncated SVD of potentially
large  matrices required by step 1 of Algorithm \ref{prune}.  Fortunately,
both needs are met by one algorithm.
